---
title: "Filter/Joins"
output: html_document
---


```{r, include=FALSE}
library (tidyverse)
library (readxl)
library (here)
```

```{r, message=FALSE}
invert_counts <- read_excel(here("r-workshop-data/invert_counts_curated.xlsx"))
fish_counts <- read_csv(here("r-workshop-data/fish_counts_curated.csv"))
kelp_counts <- read_excel(here("r-workshop-data/kelp_counts_curated.xlsx"))
```

include=FALSE does not display process or responses, message=FALSE includes process but no message, 

- select () function: keep, exclude, or rearrange columns
- filter () function: keep or exclude ROWS (observations) based on whether or not they match variable conditions

### filter to match rows based on a character string

Only keep observations from fish_counts where the common name variable matches "garibaldi"

```{r}
fish_garibaldi <- fish_counts %>%
  filter(common_name=="garibaldi")
```

### filter based on numerical observations

```{r}
fish_over_50 <- fish_counts %>%
  filter(total_count>=50)
``` 

### only find mo

```{r}
mohawk_fish <- fish_counts %>% 
  filter (site=="mohk")
```

### filter for this or that

Use %in% operator to look for any matches within a vector

c() is to combine strings

```{r}
fish_3sp <- fish_counts %>% 
  filter (common_name %in% c("garibaldi","blacksmith","black surperch"))
```

### filter to return rows for multiple variables (AND statements)

find site is abur or mohk and species is purple urchin

can seperate AND conditions within a filter line by comma or & or can %>% to an additional filter line


```{r}
urchin_abur_mohk <- invert_counts %>%
  filter (site %in% c("abur","mohk"), common_name=="purple urchin")
```










