---
title: "Filter/Joins"
output: html_document
---


```{r, include=FALSE}
library (tidyverse)
library (readxl)
library (here)
```

```{r, message=FALSE}
invert_counts <- read_excel(here("r-workshop-data/invert_counts_curated.xlsx"))
fish_counts <- read_csv(here("r-workshop-data/fish_counts_curated.csv"))
kelp_counts_abur <- read_excel(here("r-workshop-data/kelp_counts_curated.xlsx"))
```

include=FALSE does not display process or responses, message=FALSE includes process but no message, 

- select () function: keep, exclude, or rearrange columns
- filter () function: keep or exclude ROWS (observations) based on whether or not they match variable conditions

### filter to match rows based on a character string

Only keep observations from fish_counts where the common name variable matches "garibaldi"

```{r}
fish_garibaldi <- fish_counts %>%
  filter(common_name=="garibaldi")
```

### filter based on numerical observations

```{r}
fish_over_50 <- fish_counts %>%
  filter(total_count>=50)
``` 

### only find mo

```{r}
mohawk_fish <- fish_counts %>% 
  filter (site=="mohk")
```

### filter for this or that

Use %in% operator to look for any matches within a vector

c() is to combine strings

```{r}
fish_3sp <- fish_counts %>% 
  filter (common_name %in% c("garibaldi","blacksmith","black surperch"))
```

### filter to return rows for multiple variables (AND statements)

find site is abur or mohk and species is purple urchin. == means equal to

can seperate AND conditions within a filter line by comma or & or can %>% to an additional filter line


```{r}
urchin_abur_mohk <- invert_counts %>%
  filter (site %in% c("abur","mohk"), common_name=="purple urchin")
```


create subset from fish counts with garibaldi and rock wrasse where count is below 10

```{r}
low_gb_rw <- fish_counts %>%
  filter (common_name %in% c("garibaldi","rock wrasse") & total_count < 10)
  
```

filter rows that do not match conditions
use != indicateds does not match

```{r}
no_mohk <- fish_counts %>% 
  filter (site != "mohk")
```

example of filter as part of piped sequence 

all years, all sites, total or just senorita

```{r}
senoitas <- fish_counts %>%
  filter (common_name == "senorita") %>%
  group_by(site)%>%
  summarise(count_all_years = sum(total_count))
```




write values in year column as numeric (as opposed to how it was recognized as character)
```{r}
kelp_counts_abur <- kelp_counts_abur %>%
  mutate (year = as.numeric(year))
```


### Joining data frames

- `full_join` : keeps everything (safe join)
- `left_join` : merging A and B, will only return observations from B that have matches in A

| (above enter) is also OR

```{r}
fish_2016_2017 <- fish_counts %>%
  filter (year == 2016 | year == 2017)

```

join with a new frame

```{r}
abur_kelp_join <- fish_2016_2017 %>%
  full_join(kelp_counts_abur, by = c("year","site")
```








