---
title: "Intro-to-GGPlot2"
author: "Nick Munday"
date: "Sunday December 8"
output: html_document
---

### attach packages

``` {r attach-packages, message=FALSE}

library(tidyverse)
library(readxl)
library(here)

```

### read in data

```{r read-in-excel}
ca_np <- read_csv(here("r-workshop-data","ca_np.csv"))
ci_np <- read_excel(here("r-workshop-data","ci_np.xlsx"))
```

### make a graph

1. tell R we are using `ggplot`
2. tell R what data we're using, and what variables we want to plat
3. what type of graph we are creating (geom)

```{r make-graph, eval = FALSE}
ggplot (data = THIS, aes (x= THIS, y=THIS)) + 
  geom_this()
# EVAL=FALSE DOES NOT TRY TO RUN CODE
```

```{r}
ggplot (data = ci_np, aes (x=year, y=visitors)) + 
  geom_line()
```

```{r}
ggplot (data = ci_np, aes (x=year, y=visitors)) + 
  geom_area ()
```

```{r}
ggplot (data = ci_np, aes (x=year, y=visitors)) + 
  geom_col ()
```

### customize `ggplot` graphs
 - color = (update the point or line color )
 - fill = (update the fill color)
 - linetype =
 - pch = value
 - size
 - alpha

```{r}
ggplot (data = ci_np, aes (x=year, y=visitors)) + 
  geom_col (color = "red", linetype= "dashed", fill="blue")
```

I have some awesome text and now a graph below but no code will be displayed because of echo=false

```{r, echo=FALSE}
ggplot (data = ci_np, aes (x=year, y=visitors)) + 
  geom_point (color = "blue", pch = 12, size = 4, alpha=0.5)
```

### customize elements based on variable values or outcomes

```{r}
ggplot (data = ci_np, aes (x = year, y = visitors)) +
  geom_point (aes (size = visitors, 
                   color = visitors), 
              pch = 17,
              alpha = 0.75)
```

```{r}
ggplot (data = ci_np, aes (x = year, y = visitors)) +
  geom_col (aes (fill = visitors), 
            color = "magenta")
```

```{r}
ggplot (data = ci_np, aes (x = year, y = visitors)) +
  geom_line (size=3,
             color = "red") +
  theme_light()
```

### add update labels and titles with `labs`

```{r}
ggplot (data = ci_np, aes (x = year, y = visitors)) +
  geom_point (size=2,
             color = "yellow") +
  theme_dark() +
  labs (x = "My x axis title",
        y = "Visitors baby",
        title = "hola hola hola",
        subtitle = "subtitle goes here")
```

### combine compatible `geoms`, eg point plot on top of line plot

```{r}
ggplot (data = ci_np, aes (x = year, y = visitors)) +
  geom_line (color = "purple") +
  geom_point (aes (size = year),
             color = "yellow") +
  theme_dark() +
  labs (x = "My x axis title",
        y = "Visitors baby",
        title = "hola hola hola",
        subtitle = "subtitle goes here")
```


### create line graph with multiple series

```  {r}
ggplot(data = ca_np, aes(x = year, y = visitors, color= park_name)) +
  geom_line ()
```

### faceting ggplot graphs

- each series (group) gets it's own plot panel

``` {r}
ggplot (data = ca_np, aes (x = year, y = visitors, color = park_name))+
  geom_line(show.legend = FALSE) +
  theme_minimal() +
  facet_wrap (~ park_name)
```

### use ggsave to export a plot
exports most recently run graph
```{r}
ggsave ("np_graph.png", dpi = 180, width = 8, height = 7)
```






