---
title: "Intro into readxl"
author: "Nick Munday"
date: "December 7, 2019"
output: html_document
---

# Learning `readxl`

We are working with data and it's goinmg to be amazing!

We're going to use packages:
- tidyverse
- readxl
- writexl
- here

```{r packages}
library (tidyverse)
library (readxl)
library (writexl)
library (here)

### this is an annotation in code
```
<!-- this is to hide in non-code chunks
all html tags recognized -->

### Read in fish counts

```{r read-in-csv}
###Read in the fish_counts_curated.csv file
fish_counts <- read_csv(here("r-workshop-data","fish_counts_curated.csv"))
```

### Read in data from a single excel worksheet

Now, to read in an excel spreadsheet, we'll use the function `read_excel()`
``` {r read-in-excel}
### skip first 4 rows, replace -9999 with na
substrate_cover <- read_excel(here("r-workshop-data","substrate_cover_curated.xlsx"), 
                              skip = 4, 
                              na = "-9999")

```

### Use read_excel() to read in a range of values

``` {r read-in-range}
carp_cover_2000 <- read_excel(here("r-workshop-data","substrate_cover_curated.xlsx"),
                              range = "D12:I15",
                              col_names = c("site",
                                            "transect",
                                            "quad",
                                            "side",
                                            "type",
                                            "coverage"))
### c( ) is to combine arguments
```

### read in data from different tabs

read in from sheet named golb

``` {r read-in-sheets}
kelp_golb <- read_excel(here("r-workshop-data","kelp_counts_curated.xlsx"),
                          sheet = "golb")
```

### read in and combine sheet using purrr::map_df()


``` {r combine-sheets}
#create path to data
kelp_path <- here("r-workshop-data","kelp_counts_curated.xlsx")

#kelp_all_sites <- kelp_path %>%
#  excel_sheets() %>% 
#  set_names() %>% 
#  map_df(read_excel, kelp_path)
```

### write back to excel

Write "kelp_golb" object to csv

``` {r write-to-excel}
write_csv (kelp_golb, "kelp_golb_july.csv")

write_xlsx (kelp_golb, here("r-workshop-data","july.xlsx"))

```










