---
title: "Pivot-Tables"
output: html_document
---

### attach libraries

```{r}
library (tidyverse)
library(here)
library (readxl)
library (skimr) # install.packages("skimar")
```

### read in data

```{r}
lobsters <- read_xlsx(here("r-workshop-data/lobsters.xlsx"))
```

### explore data to see structure
`head` provides first 6 lines of data source

```{r}
head(lobsters)
summary (lobsters)
```
### skimr::skim() gives additional summary stats and is pipable (%>%)

```{r}
skim(lobsters)
```
```{r}
ggplot (data=lobsters, aes(x=year)) +
  geom_bar() +
  facet_wrap(~site)
```

### task

how has the average size of lobsters changed over time across sites?

Pivot tables in R are `group_by` %>% `summarize`

```{r}
lobsters %>% #pipe %>% is like "and then..." eg take lobsters data, and then group by year, and then summarize
  group_by(year)%>%
  summarize(count_by_something = n())
```

### group by multiple variables
and summarize by multiple variables

We used mean() to calculate the mean size of lobsters by each site and year and used na.rm = TRUE to calculate mean despite having missing values in our raw data

```{r}
lobsters %>%
  group_by (site, 
            year) %>%
  summarize (count_by_siteyear = n(),
             mean_size_mm = mean(size_mm, na.rm=TRUE))
```

```{r}
siteyear_summary <- lobsters %>%
  group_by (site, 
            year) %>%
  summarize (count_by_siteyear = n(),
             stanard_deviation = sd(size_mm, na.rm = TRUE),
             mean_size_mm = mean(size_mm, na.rm=TRUE),
             median_size_mm = median(size_mm, na.rm=TRUE))

siteyear_summary
```

``` {r}
ggplot (data=siteyear_summary, aes(x=year, y=median_size_mm, color = site)) +
  geom_line () 
```


``` {r}
ggsave ("lobsters_graph.png", dpi = 180, width = 8, height = 7)
```


### mutate () adds a column
it will make a new variable

add column which is size in meters

```{r}
head(lobsters)

lobsters %>%
  mutate (size_m = size_mm / 1000, 
          observer = "Nick Munday")
```

### select () function to subset variables by column
select date, site, size_m

```{r}
lobsters_subset <- lobsters %>%
  mutate (size_m = size_mm / 1000) %>%
  select(date, site, size_m)
```


